Indexing Variation Graphs\footnote{This work was supported by the Wellcome Trust grant [098051].}

### 2016-04-07 paper/paper.tex
Started writing the section on path indexes.

We can use the sampled suffix array pointers to find the text positions containing the occurrences. If $\SA[i]$ is not sampled, we start iterating $\LF(i)$, until we find a sampled pointer. If we find a sample at $\SA[\LF^{k}(i)]$, we know that
$$
\SA[i] = (\SA[\LF^{k}(i)] + k) \bmod n.
$$
If we have sampled one out of $d$ suffix array pointers at regular intervals, finding each occurrence takes $\Oh{d}$ rank queries. If we also sample one out of $d'$ \emph{inverse suffix array} pointers\footnote{The suffix array is a permutation of $\set{0, \dotsc n-1}$, and the inverse suffix array is the inverse permutation.}, we can \emph{extract} an arbitrary substring $X$ of the text using $\Oh{\abs{X}+d'}$ rank queries. In a typical case, we expect an \FMindex{} to take less than $n \log \sigma$ bits of space and to be able to find about 100,000 pattern occurrences per second \cite{Ferragina2009a}.

@@

find the start nodes of paths with label matching the pattern

general difficulty: exponential growth
GCSA \cite{Siren2014}

k-mer indexes based on hash tables, sorted k-mers

similarities between k-mer indexes and de Bruijn graphs

>>>>

We can use the sampled suffix array pointers to find the text positions containing the occurrences. If $\SA[i]$ is not sampled, we start iterating $\LF(i)$, until we find a sampled pointer. If we find a sample at $\SA[\LF^{k}(i)]$, we know that
$$
\SA[i] = (\SA[\LF^{k}(i)] + k) \bmod n.
$$
If we have sampled one out of $d$ suffix array pointers at regular intervals, finding each occurrence takes $\Oh{d}$ rank queries. If we also sample one out of $d'$ \emph{inverse suffix array} pointers\footnote{The suffix array is a permutation of $\set{0, \dotsc, n-1}$, and the inverse suffix array is the inverse permutation.}, we can \emph{extract} an arbitrary substring $X$ of the text using $\Oh{\abs{X}+d'}$ rank queries. In a typical case, we expect an \FMindex{} to take less than $n \log \sigma$ bits of space and to be able to find about 100,000 pattern occurrences per second \cite{Ferragina2009a}.

@@

A \emph{path index} is a generalization of text indexes for \emph{labeled graphs}. Given a path index for graph $G = (V, E)$, we want to use the index to find the \emph{start nodes} $v_{0} \in V$ of the paths $P = v_{0} \dotsm v_{\abs{X}-1}$ \emph{matching} the pattern $X$ (paths $P$ with $G.\glabel(P) = X$).

The design of a path index is a trade-off between the maximum supported query length and the size of the index. In the worst case, a graph may have up to $(\sigma-2)^{k}$ distinct labels for paths of length $k$ (ignoring paths containing the source node or the sink node), and the length of possible paths is unlimited in cyclic graphs. GCSA \cite{Siren2014} tried to avoid this by considering only a relatively simple class of graphs, using \emph{pruning heuristics} to simplify the graphs when necessary. The price of this was the possibility of \emph{false negatives}: path labels that exist in the graph but not in the index.

% FIXME from here
k-mer indexes based on hash tables, sorted k-mers

similarities between k-mer indexes and de Bruijn graphs: order-k+1 index can simulate order-k DBG; order-k DBG can be used as an order-k index

define de Bruijn graphs for graphs (infinite strings, no source node, repeat sink node instead)
